\name: Deterministic and Coherent Implicits

### 2017-06-07 jfp/src/Translation.tex
Translation: fix overflow
fix typos

\begin{theorem}[Type-preserving translation]\label{thm:type:preservation} Let $e$ be an $\name$
  expression, $\rulet$ be a type, $\tenv$ a type environment and $E$ be a System F expression. If
  $\tenv \turns \relation{e}{\rulet} \leadsto E$, then $|\tenv| \turns \relation{E}{|\rulet|}$.
\end{theorem}
Here we define the translation of the type environment form $\name$ to System F as:
\begin{equation*}
\begin{array}{rcl@{\hspace{2cm}}rcl}
|\epsilon| & = & \epsilon & |\tenv,\alpha| & = & |\tenv|, \alpha \\
|\tenv,x : \rulet| & = & |\tenv|, x : |\rulet| &
|\tenv, \rulet \leadsto x| & = & |\tenv|, x : |\rulet|
\end{array}
\end{equation*}

@@

Moreover, we can express three key properties of Figure~\ref{fig:resolution2}'s
definition of resolution in terms of the generated evidence.
Firstly, the deterministic version of resolution is a sound variation on the original ambiguous resolution.
\begin{lemma}[Soundness]
Figure~\ref{fig:resolution2}'s definition of resolution is sound (but
incomplete) with respect to Figure~\ref{fig:resolution1}'s definition.
\[\forall \tenv, \rulet, E: \quad\quad \tenv \ivturns \rulet \leadsto E \quad\Rightarrow\quad \tenv \vturns \rulet \leadsto E \]
\end{lemma}
Secondly, the deterministic resolution guarantees a strong form of coherence:
\begin{lemma}[Determinacy]
The generated evidence of resolution is uniquely determined.
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad \tenv \ivturns \rulet \leadsto E_1 ~\wedge~ \tenv \ivturns \rulet \leadsto E_2 \quad\Rightarrow\quad E_1 = E_2 \]
\end{lemma}
Thirdly, on top of the immediate coherence of deterministic resolution, 
an additional coherence property holds.  
\begin{lemma}[Stability]
Resolution is stable under substitution.
\[\forall \tenv,\alpha,\tenv',\sigma,\rulet, E: \quad\quad 
\tenv,\alpha,\tenv' \ivturns \rulet \leadsto E \quad\wedge\quad \tenv \vdash \sigma
\quad\Rightarrow\quad 
\tenv,\tenv'[\sigma/\alpha] \ivturns \rulet[\sigma/\alpha] \leadsto E[|\sigma|/\alpha] \]
\end{lemma}

>>>>

\begin{theorem}[Type-Preserving Translation]\label{thm:type:preservation} Let $e$ be a $\name$
  expression, $\rulet$ be a type, $\tenv$ be a type environment and $E$ be a System F expression. If
  $\tenv \turns \relation{e}{\rulet} \leadsto E$, then $|\tenv| \turns \relation{E}{|\rulet|}$.
\end{theorem}
Here we define the translation of the type environment form $\name$ to System F as:
\begin{equation*}
\begin{array}{rcl@{\hspace{2cm}}rcl}
|\epsilon| & = & \epsilon & |\tenv,\alpha| & = & |\tenv|, \alpha \\
|\tenv,x : \rulet| & = & |\tenv|, x : |\rulet| &
|\tenv, \rulet \leadsto x| & = & |\tenv|, x : |\rulet|
\end{array}
\end{equation*}

@@

Moreover, we can express three key properties of Figure~\ref{fig:resolution2}'s
definition of resolution in terms of the generated evidence.
Firstly, the deterministic version of resolution is a sound variation on the original ambiguous resolution.
\begin{lemma}[Soundness]
Figure~\ref{fig:resolution2}'s definition of resolution is sound (but
incomplete) with respect to Figure~\ref{fig:resolution1}'s definition.
\[\forall \tenv, \rulet, E: \quad\quad \tenv \ivturns \rulet \leadsto E \quad\Rightarrow\quad \tenv \vturns \rulet \leadsto E \]
\end{lemma}
Secondly, the deterministic resolution guarantees a strong form of coherence:
\begin{lemma}[Determinacy]
The generated evidence of resolution is uniquely determined.
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad \tenv \ivturns \rulet \leadsto E_1 ~\wedge~ \tenv \ivturns \rulet \leadsto E_2 \quad\Rightarrow\quad E_1 = E_2 \]
\end{lemma}
Thirdly, on top of the immediate coherence of deterministic resolution, 
an additional coherence property holds.  
\begin{lemma}[Stability]
Resolution is stable under substitution.
\[\forall \tenv,\alpha,\tenv',\sigma,\rulet, E: \enskip\
\tenv,\alpha,\tenv' \ivturns \rulet \leadsto E \enskip\wedge\enskip \tenv \vdash \sigma
\enskip\Rightarrow\enskip 
\tenv,\tenv'[\sigma/\alpha] \ivturns \rulet[\sigma/\alpha] \leadsto E[|\sigma|/\alpha] \]
\end{lemma}
